---
title: "Analyzing r/wallstreetbets Sentiment"
author: "Oliver Hammond, Jake Reiners, Marit McQuaig"
date: "12/15/2021"
output: html_document
---

### Overview

[WallStreetBets](https://www.reddit.com/r/wallstreetbets/), a subreddit on the website “Reddit,” is a casual forum to discuss the stock market with other amateur investors. This subreddit has had a minor impact on the stock market as well; the beginning of the [GameStop short squeeze](https://www.google.com/url?q=https://archive.ph/20210130144720/https://www.wsj.com/articles/the-real-force-driving-the-gamestop-amc-blackberry-revolution-11611965586&sa=D&source=docs&ust=1638811083813000&usg=AOvVaw3g-nI4WeAOEJJgYNKBot75) in January 2021 has been attributed to users of r/wallstreetbets. In our project we investigate if there is relationship between the average daily sentiment of the WallStreetBets subreddit and the index fund \$SPY (a fund that mirrors the 500 stocks on the S&P 500). Unfortunately, we did not find any correlation between the daily changes of SPY and the average sentiment of r/wallstreetbets.

### Dataset

We used two datasets, both formatted similarly. They can be found at https://www.kaggle.com/mattpodolak/reddit-wallstreetbets-comments and https://www.kaggle.com/mattpodolak/rwallstreetbets-posts-and-comments. The first dataset contains 34 million rows: comments from 1/31/12 to 2/16/21. Around one third of the original dataset was corrupted to the point that we couldn’t read it in, so we used the second dataset, which spans from December 2020 to February 2021 to partially fill in that gap. We focused on two variables for our analysis:

**Body:** Text of each reddit comment

**Created UTC:** Date-time of posting since the epoch of Linux-Bash in seconds

Our SPY historical values came from [The Wall Street Journal](https://www.wsj.com/).

### Preprocessing and Parallelization

We needed to use the CHTC to run parallel jobs due to the size of the dataset. We split the first dataset into 34 files of 1 million lines and the second dataset into 10 files of 1 million lines. Due to parsing issues with the original dataset, 11 chunks (~11 million comments) of the first dataset were unusable. We conducted the data splitting in Python due to it having a robust csv reader. In total we ran 33 jobs on the CHTC, each requesting 6GB of memory and 1.5GB of disk space and each took around 30 minutes to run. Most jobs require under 6GBs of memory but when requesting less one job was going over the allocated memory.

### Computation

The R library sentimentr was utilized to find the average sentiment of comments each day. Each comment was scored from -1 to 1, ranging from strongly negative to strongly positive. Since comments from the same day are spread across multiple files, we kept track of not only the average sentiment for each day but the number of times that date appeared in each dataset in order to calculate a weighted average. After processing on the CHTC, we aggregated the results and calculated the average sentiment for each day.

### Analysis
Once the computations completed, we graphed the daily average sentiment against the daily changes in the SPY stock to look for a relationship. The average allows easy comparison between days because the total number of comments will not affect values of the mean and the average will always fall between -1 and 1. 

We also compared the cumulative sentiment to the daily closing values of the S&P to see if the community at r/wallstreetbets picked up on more general trends. The sum allows us to look at the subreddit from a different angle: eventful days in the stock market may be reflected by much more extreme values.

### Results

When comparing the average cumulative sentiment of the subreddit with the SPY closing value, there is a slight correlation (correlation coefficient of 0.514). Clearly, r/wallstreetbets users picked up on major peaks and drops in the market.

```{r, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(ggpubr)

hist.spy.df <- read.csv("HistoricalPrices.csv")
hist.spy.df$Date <- as.Date(hist.spy.df$Date, "%m/%d/%y")

hist.spy.df<-hist.spy.df[nrow(hist.spy.df):1,]


close.diff <- diff(as.matrix(hist.spy.df$Close))
close.diff <- rbind(0,close.diff)
hist.spy.df$differential <- close.diff[,1]

wsb.sent.df <- read.csv("combined_sentiment.csv")
wsb.sent.df$total_sent <- wsb.sent.df$ave_sentiment * wsb.sent.df$day_count
wsb.agg.df <- aggregate(cbind(total_sent, day_count) ~ day, wsb.sent.df, FUN=sum)
wsb.agg.df$day <- as.Date(wsb.agg.df$day)
wsb.agg.df$avg_sent <- wsb.agg.df$total_sent/ wsb.agg.df$day_count

wsb.agg.df <- wsb.agg.df[order(wsb.agg.df$day), ]
combined.df <- merge(wsb.agg.df,hist.spy.df, by.x="day",by.y="Date")

# Calculate Cumulative Sentiment
combined.df$sent_cumsum <- cumsum(combined.df$total_sent)

# Calculate cumulative average sentiment
combined.df$avgsent_cumsum <- cumsum(combined.df$avg_sent)

ggplot(combined.df, aes(x=day)) +
  geom_line(aes(y=avgsent_cumsum, color='Avg Sent Sum')) +
  geom_line(aes(y=Close / 129, color='SPY Close')) + 
  scale_y_continuous(name="r/wallstreebets Average Cumulative Sentiment", sec.axis = sec_axis(~.*129,name="SPY $ Close")) +
  xlab("Date") +
  scale_color_manual("Plot Lines",values = c("Avg Sent Sum"="red2","SPY Close"="cyan3")) + 
  ggtitle("Cumulative Average Sentiment of WSB with SPY Closing Value") +
  theme(legend.position = c(0.8,0.2))

```

Due to the parsing issues, we had some gaps over dates in our dataset but were able to compare data at four different time periods. We found no strong correlation between the r/wallstreetbets comments and the S&P daily differential (correlation coefficients below 0.20).

```{r, echo=FALSE, results=FALSE, message=FALSE, warning=FALSE}
plot1 <- ggplot(combined.df, aes(x=day)) +
  geom_line(aes(y=avg_sent, color='Sentiment'), alpha = .6) +
  geom_line(aes(y=differential / 290, color='Differential'), alpha = .6) + 
  scale_y_continuous(name="Average Sentiment", sec.axis = sec_axis(~.*290,name="SPY $ Differential")) +
  xlab("Date") +
  scale_color_manual("Plot Lines",values = c("Sentiment"="red2","Differential"="blue2")) +
  scale_x_date(date_breaks = "3 month",
               limits = as.Date(c('2020-07-01','2021-02-16'))) +
  theme(legend.position = c(0.12,0.85))
  
 
plot2 <- ggplot(combined.df, aes(x=day)) +
  geom_line(aes(y=avg_sent, color='Sentiment'), alpha = .6) +
  geom_line(aes(y=differential / 290, color='Differential'), alpha = .6) + 
  scale_y_continuous(name="Average Sentiment", sec.axis = sec_axis(~.*290,name="SPY $ Differential")) +
  xlab("Date") +
  scale_color_manual("Plot Lines",values = c("Sentiment"="red2","Differential"="blue2")) +
  scale_x_date(date_breaks = "2 month",
               limits = as.Date(c('2020-01-01','2020-07-01'))) +
  theme(legend.position = c(0.12,0.85))


plot3 <- ggplot(combined.df, aes(x=day)) +
  geom_line(aes(y=avg_sent, color='Sentiment'), alpha = .6) +
  geom_line(aes(y=differential / 290, color='Differential'), alpha = .6) + 
  scale_y_continuous(name="Average Sentiment", sec.axis = sec_axis(~.*290,name="SPY $ Differential")) +
  xlab("Date") +
  scale_color_manual("Plot Lines",values = c("Sentiment"="red2","Differential"="blue2")) +
  scale_x_date(date_breaks = "2 month",
               limits = as.Date(c('2019-07-01','2020-01-01'))) +
  theme(legend.position = c(0.12,0.85))


plot4 <- ggplot(combined.df, aes(x=day)) +
  geom_line(aes(y=avg_sent, color='Sentiment'), alpha = .6) +
  geom_line(aes(y=differential / 290, color='Differential'), alpha = .6) + 
  scale_y_continuous(name="Average Sentiment", sec.axis = sec_axis(~.*290,name="SPY $ Differential")) +
  xlab("Date") +
  scale_color_manual("Plot Lines",values = c("Sentiment"="red2","Differential"="blue2")) +
  scale_x_date(date_breaks = "2 month",
               limits = as.Date(c('2019-02-19','2019-07-01'))) +
  theme(legend.position = c(0.12,0.85))

plot <- ggarrange(plot4,plot3,plot2,plot1,ncol=2,nrow = 2,common.legend=TRUE, legend="bottom")
annotate_figure(plot, top = text_grob("Average Daily WSB Sentiment Compared to SPY Daily Change at Four Time Periods"))
```

### Conclusion

Our analysis doesn't point to any meaningful relationship between r/wsb sentiment and the stock market. This may simply be because reddit is an open forum on the internet where anyone can say anything they want. The abundance of sarcasm, emojis, and other slang in the comments also could've made sentiment analysis performed by a machine unreliable. Future work to make this analysis stronger could include conducting sentiment analysis using a homemade dictionary with slang (and associated sentiment) specific to this subreddit.


#### [GitHub Link](https://github.com/ohammond1/605FinalProject)




